---
title: "Group 1 — Clinic bodies"
subtitle: "Studio wall"
author: "Dheemant, Madhumitha, Sadiyah, Vidhi"
format:
  html:
    toc: true
    toc-depth: 2
---

# Spine

> Who is well in the clinic, who is labelled ill, and where does infant death still sit on the map?
Both panels together show who in the clinic is labelled ill: how heart disease splits across sex and chest-pain type, and whether high blood sugar and exercise-induced angina go together more than chance would predict.

# Introduction — the tables we own

- **Panel A** (`panel-A.csv`) — clinic measurements: age, blood pressure, cholesterol, maximum heart rate, plus sex and heart-disease label for grouping.
- **Panel B** (`panel-B.csv`) — the *same people*, Quals only: chest pain, blood sugar flag, exercise angina, heart disease.
- **Panel C** (`panel-C.csv`) — infant mortality by country. No geometry in the file; we join to a world map in R.

A does not grow a mosaic. B does not grow a honest histogram of chest pain. C is a rate, not a headcount.

We load from this folder:

Found tables train the mark. A campus Free Hunch, if we collect one, is a **made** panel of the same pair type — new people, old grammar. It does not replace A, B, or C.

```{r}
#| label: Set up chunk
library(tidyverse)
library(skimr)
library(mosaic)
library(ggformula)
library(vcd)
```

# Panel A — body / measure

*Age, BP, Cholesterol, MaximumHR are Quant. Sex and HeartDisease are Quals used to *split* the Quants.*

```{r}
#| label: My Panel A data
A <-  read_csv(file = "data/panel-A.csv")
```

### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

The Qualitative variable in this table entail - HeartDisease and Sex(These were character variables which were mutated to factor variables).

the quantitative variables in this table are - Age, BP, Cholestrol, MaximumHR.

A few piecies of data missing here are the fatality rate and the fact that casuality cannot be proved since this is not an RCT data set.

```r
# glimpse() / inspect() / skim()
```

```{r}
#| label: Table for Panel A
glimpse(A)
skim(A)
```

```{r}
#| label: change to factor(Panel A)
A |> mutate(Sex = as.factor(Sex))
A |> mutate(HeartDisease = as.factor(HeartDisease))
A<- A |> dplyr::mutate(Sex = as.factor(Sex), HeartDisease = as.factor(HeartDisease))
glimpse(A)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

Question can answer:
1.How does cholestrol affect different sex?
2.Based on sex and age how does the maximum Heartrate fluctuate?
3.How does the age distribution of people with and without heart disease differ in this dataset?
4.how does the number of men and women having BP vary over the years?

Questions it cannot answer:

1.For what % of the population is cholestrol the reason for heart disease?
2.Does sex itself influence a person's maximum heart rate, or are the differences in maximum heart rate actually related to other factors such as age or health?

### 3. Choose a mark

Questions that can be answered:

```{r}
#| label: Chart1-Question1A
A |> gf_boxplot(~ Cholesterol, bins = 50, fill = ~ Sex)
```

```{r}
#| label: Chart2-Question2A
A |> gf_point(MaximumHR~Age, color = ~Sex) |>  gf_smooth(method = "lm") |>  gf_density_2d()
```

```{r}
#| label: Chart3-Question3A
A |> gf_boxplot(Age ~ HeartDisease)
```

```{r}
#| label: Chart4-Question4A
A |> gf_line(BP ~ Age,bins= 100, color =  ~Sex)
```

Questions that cannot be answered:
```{r}
#| label: Chart1-Question1AN
A |> gf_point(Cholesterol~Sex,color = ~HeartDisease)
```

```{r}
#| label: Chart2-Question2AN
A |> ggplot(aes(Sex, MaximumHR, fill = Sex)) + geom_violin(alpha = 0.6) + geom_jitter(width = 0.08, alpha = 0.3, size = 1) + theme_minimal()
```

Statistic Tests:

```{r}
#| label: StatisticTest1-Summarize
A |> group_by(Sex) |> summarize( mean_MaximumHR = mean(MaximumHR, na.rm = TRUE))
A |> group_by(Sex) |> summarize( mean_Cholesterol = mean(Cholesterol, na.rm = TRUE))
A |> group_by(Sex) |> summarize( mean_BP = mean(BP, na.rm = TRUE))
```

```{r}
#| label: StatisticTest2-T-Test
mosaic::t_test(MaximumHR ~ Sex, data = A)
mosaic::t_test(Cholesterol ~ Sex, data = A)
mosaic::t_test(BP ~ Sex, data = A)
```

### 4. Predict: What do we think we will see?

We predict that age, sex, blood pressure, and cholesterol will show noticeable differences in maximum heart rate and heart disease.

### 5. Say

Title:
Age and sex show different patterns in heart health measures.

Caption:
The graphs show how age, sex, blood pressure, cholesterol, maximum heart rate, and heart disease are distributed and related across the sample.

One sentence we will stand behind:
The data shows differences between groups, but these patterns do not prove that one factor causes heart disease.

### 6. Limit
What this panel cannot do. Which kit table we refused to abuse.

The data available to us in this panel cannot prove dependency in terms of causality between two or more variables.

# Panel B — category / talk

*ChestPain, Sex, HeartDisease, ExerciseInducedAngina. Mosaic or stacked bar. No mean of ChestPain.*

```{r}
#| label: My Panel B data
B <- read_csv(file = "data/panel-B.csv")
```

### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```

```{r}
#| label: Table For Panel B
glimpse(B)
skim(B)
```

```{r}
#| label: Change to factor(Panel B)
B <- B |> mutate(Sex = as.factor(Sex))
B <- B |> mutate(across(where(is.character),as.factor))
B |> glimpse()
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

Question can answer:
1. What part of the population posses both blood sugar and Heart Disease?
2. How does the proportion of people with heart disease differ across different chest pain types?
3. How does the distribution of exercise-induced angina differ between males and females?

Questions it cannot answer:
1.How often does exercise-induced angina cause chest pain among man and women?
2.Why do certain chest pain types have higher rates of heart disease?
3.What factors are responsible for certain people having heart disease?

### 3. Choose a mark

Questions that can be answered:
```{r}
#| label: Chart1-Question1B
B |> ggplot(aes(x = BloodSugar, fill = HeartDisease)) +geom_bar() +
  labs(x = "Blood Sugar", y = "Number of People")
```

```{r}
#| label: Chart2-Question2B
B |> ggplot(aes(x = ChestPain, fill = HeartDisease)) + geom_bar(position = "fill") +labs(title = "Heart Disease Proportion Across Chest Pain Types",x = "Chest Pain Type",y = "Proportion",fill = "Heart Disease") + theme_classic()
```

```{r}
#| label: Chart3-Question3B
B |> 
  ggplot(aes(x = ExerciseInducedAngina, fill = HeartDisease)) + geom_bar(position = "dodge") +facet_wrap(~Sex) +
  labs(x = "Exercise-Induced Angina", y = "Number of People")
```

Questions that cannot be answered:

```{r}
#| label: Chart1-Question1BN
vcd::structable(ChestPain ~ ExerciseInducedAngina + Sex, data = B) |> 
  vcd::mosaic(labeling = labeling_values, shade = TRUE, type = "expected")
```

```{r}
#| label: Chart2-Question2BN
B |> 
  ggplot(aes(x = ChestPain, fill = HeartDisease)) + geom_bar() + coord_flip() +
  labs(x = "Chest Pain", y = "Number of People")
```


```{r}
#| label: Chart3-Question3BN
vcd::structable(ExerciseInducedAngina ~ BloodSugar, data = B) |> 
  vcd::mosaic(shade = TRUE, gp = shading_Friendly)
```

Statistic Test

```{r}
#| label: Chi-square_test
mosaic::xchisq.test(BloodSugar ~ ExerciseInducedAngina, data = B)
mosaic::xchisq.test(ChestPain ~ ExerciseInducedAngina, data = B)
```

Observation and conclusions drawn:

Blood Sugar ~ Exercise Induced Angina-
The above test proves that there is no significant correlation between the two variables. It states that the Blood Sugar and Exercise Induced Angina are independent of each other.
Further since the p value exceeds 0.5 it fails to prove the fact that nothing is happening.

Chest Pain ~ Exercise Induced Angina-
The above test proves that there is strong association between the two variables. Chest Pain and Exercise Induced Angina are Highly dependent on each other.
The low p value proves the hypothesis that nothing is happening.


### 4. Predict:What do we think we will see? One sentence.

We predict that heart disease will be more common among people with certain chest pain types, exercise-induced angina, and high blood sugar.

### 5. Say

Title:
Heart disease shows different patterns across several health-related factors.

Caption:
The graphs compare heart disease with chest pain, blood sugar, exercise-induced angina, and sex to identify patterns within the sample.

One sentence we will stand behind:
The data shows associations between heart disease and these factors, but it cannot tell us what causes heart disease.

### 6. Limit

What this panel cannot do:
This panel can show associations between the variables, but it cannot explain why these patterns occur or prove that one factor causes heart disease.

Which kit table we refused to abuse:
The contingency table.


# Panel C — place

*Join `country` to `rnaturalearth` / `tmap::World`. Fill is infant mortality (a rate).*

Geometry is not in the CSV. Join in R, then use the same seven headings.

```r
# library(sf)
# library(rnaturalearth)
# world <- ne_countries(scale = "medium", returnclass = "sf")
# # join on country name; list the unmatched labels
```


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```
```{r}
myPanelCData <- read_csv(file = "data/panel-C.csv")
glimpse(myPanelCData)
```

```{r}

myPanelCData <- read_csv(file = "data/panel-C.csv")
skim(myPanelCData)

```

```{r}

myPanelCData <- read_csv(file = "data/panel-C.csv")

inspect(myPanelCData)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Conclusion

What the spine looks like after three panels.

- The question we will defend in ninety seconds
- What we will **not** claim
- Which lines were human, which were machine
- Limits: sample, missingness, the test we did not run
- Gift for the steal-this strip (palette, annotation, join trick)

```r
# sessionInfo()
```